A Note on Convergence Diagnostics for Multiple Imputation using Chained Equations (MICE)

### 2019-12-11 Simulation/Writeup/NoteConvergence.Rnw
Some feedback from Gerko

The goal of this research project is to develop novel methodology and guidelines for evaluating MI methods. These evaluation measures and guidelines will subsequently be implemented in an interactive evaluation framework for multiple imputation to aid applied researchers in drawing valid inference from incomplete datasets. This note provides the theoretical foundation for one of these evaluation measures that is vital in evaluating MI procedures: a diagnostic to assess convergence of the MI algorithm. More specifically, this note focuses on the MI algorithm that is implemented in the \proglang{R} package \pkg{mice}: `Multiple Imputation using Chained Equations' (MICE). The research question that will be addressed is: %. a measure that is vital in evaluating one vital measure for evaluating MI procedures--an algorithmic convergence diagnostic. This note specifically investigates one vital evaluation measure: how to diagnose convergence of a multiple imputation algorithm. 
%This note focusses on one measure that is vital for evaluating the validity of multiply imputed data: a convergence diagnostic. of the MI algorithm. 
%Therefore, the aim of this research project is to investigate which measures are vital for . how the convergence properties of the MI algorithm that is implemented in \pkg{mice}: `Multiple Imputation using Chained Equations' (MICE). This note addresses the question: 
\emph{`How to diagnose convergence of the multiple imputation algorithm MICE?'.} \textbf{REMOVE THE :'s HERE!}

@@

To investigate the convergence properties of the \pkg{mice} algorithm, different lengths of imputation chains are considered. This simulation study considers 100 simulation conditions. Each condition has a different maximum number of iterations, varying between one and one-hundred iterations. In this study, 1000 simulation runs are performed for each of the 100 simulation conditions. \textbf{Estimands are convergence diagnostics. Performance measures are simulation diagnostics.  where for one through one hundred the Rhat and ... are calculated.} The simulation set-up consists of several steps, summarized in the pseudo-code below. The complete \proglang{R} script of the simulation study is available on Github.

>>>>

The goal of this research project is to develop novel methodology and guidelines for evaluating MI methods. These evaluation measures and guidelines will subsequently be implemented in an interactive evaluation framework for multiple imputation to aid applied researchers in drawing valid inference from incomplete datasets. This note provides the theoretical foundation towards the diagnostic evaluation of the convergence of MI algorithms. % for diagnostics for evalua one of these evaluation measures that is vital in evaluating MI procedures: a diagnostic to assess convergence of the MI algorithm. 
For reasons of brevity, this note only focuses ion the MI algo implemented in mice. %The implementation limits itself to the MI algorithm 
%More specifically, this note focuses on the MI algorithm that is implemented in the \proglang{R} package \pkg{mice}: `Multiple Imputation using Chained Equations' (MICE) \textbf{[can be removed, can be found in the mice ref]}. 
To this end we will evaluate how convergence of the imputation algo can be diagnosed.

%The research question that will be addressed is: %. a measure that is vital in evaluating one vital measure for evaluating MI procedures--an algorithmic convergence diagnostic. This note specifically investigates one vital evaluation measure: how to diagnose convergence of a multiple imputation algorithm. 
%This note focusses on one measure that is vital for evaluating the validity of multiply imputed data: a convergence diagnostic. of the MI algorithm. 
%Therefore, the aim of this research project is to investigate which measures are vital for . how the convergence properties of the MI algorithm that is implemented in \pkg{mice}: `Multiple Imputation using Chained Equations' (MICE). This note addresses the question: 
%\emph{`How to diagnose convergence of the multiple imputation algorithm MICE?'.}

@@

To investigate the convergence properties of the \pkg{mice} algorithm, we perform model-based simulation of the MI procedure with varying imputation chain lengths. Each simulation condition has a different maximum number of iterations, from one to one-hundred iterations. We compute $\widehat{R}$ and auto-correlation (`convergence diagnostics') and performance measures (`simulation diagnostics') for each condition. The number of simulation runs per condition is 1000. %All simulation conditions are replicated 1000 times. %For each of the 100 conditions 1000 simulation runs are performed. for each of the 100 simulation conditions. \textbf{Estimands are convergence diagnostics. Performance measures are simulation diagnostics.  where for one through one hundred the Rhat and ... are calculated.} 
The simulation set-up consists of several steps, summarized in the pseudo-code below.\footnote{The complete \proglang{R} script of the simulation study is available on Github.}
